---
title: "Compare MCL1 Across Screens Figures"
author: "Neekesh Dharia & Guillaume Kugener"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, include=FALSE, message=FALSE, warning = FALSE)
source('setup.R')
source('load_data.R')
```


# Generated using: `r toupper(version_to_use)`


```{r load_needed_datasets}
# This script uses the following datasets:
mf <- mf
gene_dependency <- load_gene_dependency()
gene_effect <- load_gene_effect()
gene_expression <- load_gene_expression()
gene_lrt_table <- load_lrt_table()
achilles_common_essentials <- load_achilles_common_essentials()

if(!dir.exists('figures')) { dir.create('figures') }
if(!dir.exists('figures/mcl1_compare')) { dir.create('figures/mcl1_compare') }
```

```{r}
load_sanger_gene_effect <- function(){
  if(!file.exists("data/sanger_gene_effect.csv")) # If file does not exist, download from FigShare. Otherwise load local file.
  {
    sanger_gene_effect <- fread("https://ndownloader.figshare.com/files/16623881")
    fwrite(sanger_gene_effect, "data/sanger_gene_effect.csv")
  } else
  {
    sanger_gene_effect <- fread("data/sanger_gene_effect.csv")
  }
  # Reformat as a matrix.
  sanger_gene_effect %<>% dplyr::filter(V1 %in% mf$DepMap_ID)
  rownames_temp <- sanger_gene_effect$V1
  sanger_gene_effect <- as.matrix(sanger_gene_effect[,-1])
  rownames(sanger_gene_effect) <- rownames_temp
  return(sanger_gene_effect)
}

load_demeter_gene_effect <- function(){
  if(!file.exists("data/demeter_gene_effect.csv")) # If file does not exist, download from FigShare. Otherwise load local file.
  {
    demeter_gene_effect <- fread("https://ndownloader.figshare.com/files/13515395")
    fwrite(demeter_gene_effect, "data/demeter_gene_effect.csv")
  } else
  {
    demeter_gene_effect <- fread("data/demeter_gene_effect.csv")
  }
  # Reformat as a matrix.
  rownames_temp <- demeter_gene_effect$V1
  demeter_gene_effect <- as.matrix(demeter_gene_effect[,-1])
  rownames(demeter_gene_effect) <- rownames_temp
  demeter_gene_effect <- t(demeter_gene_effect)
  demeter_gene_effect <- demeter_gene_effect[rownames(demeter_gene_effect) %in% mf$CCLE_name,]
  ccle_to_depmap <- mf %>% select(DepMap_ID, CCLE_name)
  rownames(ccle_to_depmap) <- ccle_to_depmap$CCLE_name
  rownames(demeter_gene_effect) <- ccle_to_depmap[rownames(demeter_gene_effect), "DepMap_ID"]
  return(demeter_gene_effect)
}

load_demeter_drive_gene_effect <- function(){
  if(!file.exists("data/demeter_drive_gene_effect.csv")) # If file does not exist, download from FigShare. Otherwise load local file.
  {
    demeter_drive_gene_effect <- fread("https://ndownloader.figshare.com/files/11489693")
    fwrite(demeter_drive_gene_effect, "data/demeter_drive_gene_effect.csv")
  } else
  {
    demeter_drive_gene_effect <- fread("data/demeter_drive_gene_effect.csv")
  }
  # Reformat as a matrix.
  rownames_temp <- demeter_drive_gene_effect$V1
  demeter_drive_gene_effect <- as.matrix(demeter_drive_gene_effect[,-1])
  rownames(demeter_drive_gene_effect) <- rownames_temp
  demeter_drive_gene_effect <- t(demeter_drive_gene_effect)
  demeter_drive_gene_effect <- demeter_drive_gene_effect[rownames(demeter_drive_gene_effect) %in% mf$CCLE_name,]
  ccle_to_depmap <- mf %>% select(DepMap_ID, CCLE_name)
  rownames(ccle_to_depmap) <- ccle_to_depmap$CCLE_name
  rownames(demeter_drive_gene_effect) <- ccle_to_depmap[rownames(demeter_drive_gene_effect), "DepMap_ID"]
  return(demeter_drive_gene_effect)
}

load_sanger_dose_response <- function(){
  if(!file.exists("data/sanger_dose_response.csv")) # If file does not exist, download from FigShare. Otherwise load local file.
  {
    sanger_dose_response <- fread("https://depmap.org/portal/download/api/download/external?file_name=processed_portal_downloads%2Fgdsc-drug-set-export-658c.5%2Fsanger-dose-response.csv")
    fwrite(sanger_dose_response, "data/sanger_dose_response.csv")
  } else
  {
    sanger_dose_response <- fread("data/sanger_dose_response.csv")
  }
  # Reformat as a matrix.
  sanger_dose_response %<>% dplyr::filter(V1 %in% mf$DepMap_ID)
  rownames_temp <- sanger_dose_response$V1
  sanger_dose_response <- as.matrix(sanger_dose_response[,-1])
  rownames(sanger_dose_response) <- rownames_temp
  return(sanger_dose_response)
}

load_sanger_dose_response <- function(){
  if(!file.exists("data/sanger_dose_response.csv")) # If file does not exist, download from FigShare. Otherwise load local file.
  {
    sanger_dose_response <- fread("https://depmap.org/portal/download/api/download/external?file_name=processed_portal_downloads%2Fgdsc-drug-set-export-658c.5%2Fsanger-dose-response.csv")
    fwrite(sanger_dose_response, "data/sanger_dose_response.csv")
  } else
  {
    sanger_dose_response <- fread("data/sanger_dose_response.csv")
  }
  sanger_dose_response %<>% dplyr::filter(ARXSPAN_ID %in% mf$DepMap_ID)
  return(sanger_dose_response)
}

load_prism_secondary_response <- function(){
  if(!file.exists("data/prism_secondary_response.csv")) # If file does not exist, download from FigShare. Otherwise load local file.
  {
    prism_secondary_response <- fread("https://ndownloader.figshare.com/files/20237739")
    fwrite(prism_secondary_response, "data/prism_secondary_response.csv")
  } else
  {
    prism_secondary_response <- fread("data/prism_secondary_response.csv")
  }
  prism_secondary_response %<>% dplyr::filter(depmap_id %in% mf$DepMap_ID)
  return(prism_secondary_response)
}

sanger_gene_effect <- load_sanger_gene_effect()
demeter_gene_effect <- load_demeter_gene_effect()
demeter_drive_gene_effect <- load_demeter_drive_gene_effect()
sanger_dose_response <- load_sanger_dose_response()
prism_secondary_response <- load_prism_secondary_response()
```

The table below highlights all of the cell lines that are included in the analyses and the types that they are annotated as:

```{r}
cell_lines_included_table <- rbind(
  data.frame(name=row.names(gene_dependency), stringsAsFactors = F) %>% mutate(Source='Dependency'),
  data.frame(name=row.names(gene_expression), stringsAsFactors = F) %>% mutate(Source='Expression'),
  data.frame(name=row.names(sanger_gene_effect), stringsAsFactors = F) %>% mutate(Source='Sanger CRISPR'),
  data.frame(name=row.names(demeter_gene_effect), stringsAsFactors = F) %>% mutate(Source='DEMETER2 Combined RNAi'),
  data.frame(name=row.names(demeter_drive_gene_effect), stringsAsFactors = F) %>% mutate(Source='DEMETER2 DRIVE RNAi'),
  data.frame(name=unique(sanger_dose_response$ARXSPAN_ID), stringsAsFactors = F) %>% mutate(Source='GDSC'),
  data.frame(name=unique(prism_secondary_response$depmap_id), stringsAsFactors = F) %>% mutate(Source='PRISM')
) %>% mutate(value=1) %>% mutate(DepMap_ID=name) %>%
  spread(., Source, value) %>%
  left_join(., mf %>% dplyr::select(DepMap_ID, CCLE_name, Type), by='DepMap_ID') %>%
  dplyr::select(DepMap_ID, CCLE_name, everything())
```

```{r include=TRUE} 
cell_lines_included_table %>% datatable(options=list(scrollX=T))
```

# Overview

* Compare MCL1 dependency across RNAi and other CRISPR screens
* Compare MCL1 dependency to drug data from CTD2, GDSC, PRISM

```{r}
genes_to_highlight <- "MCL1 (4170)"

data_to_plot <- left_join(
  left_join(mf, 
            as.data.frame(cbind(rownames(gene_effect), gene_effect[,genes_to_highlight,drop=F])) %>% 
              set_colnames(c("DepMap_ID", "gene_effect")), by = "DepMap_ID"),
  as.data.frame(cbind(rownames(demeter_gene_effect), demeter_gene_effect[,genes_to_highlight,drop=F])) %>% 
    set_colnames(c("DepMap_ID", "demeter_gene_effect")), by = "DepMap_ID") %>%
  dplyr::mutate(gene_effect = as.numeric(gene_effect), 
                demeter_gene_effect = as.numeric(demeter_gene_effect))

data_to_plot %<>% left_join(., as.data.frame(cbind(rownames(demeter_drive_gene_effect), demeter_drive_gene_effect[,genes_to_highlight,drop=F])) %>% 
                              set_colnames(c("DepMap_ID", "demeter_drive_gene_effect")), by = "DepMap_ID") %>%
  dplyr::mutate(demeter_drive_gene_effect = as.numeric(demeter_drive_gene_effect))

data_to_plot %<>% left_join(., as.data.frame(cbind(rownames(sanger_gene_effect), sanger_gene_effect[,genes_to_highlight,drop=F])) %>% 
                              set_colnames(c("DepMap_ID", "sanger_gene_effect")), by = "DepMap_ID") %>%
  dplyr::mutate(sanger_gene_effect = as.numeric(sanger_gene_effect))

data_to_plot %<>% left_join(., prism_secondary_response %>% 
                              dplyr::filter(name == "TW-37") %>%
                              dplyr::select(depmap_id, auc) %>%
                              magrittr::set_colnames(c("DepMap_ID", "prism_auc_TW37")), by = "DepMap_ID") %>%
  dplyr::mutate(prism_auc_TW37 = as.numeric(prism_auc_TW37))

data_to_plot %<>% left_join(., prism_secondary_response %>% 
                              dplyr::filter(name == "gossypol") %>%
                              dplyr::select(depmap_id, auc) %>%
                              magrittr::set_colnames(c("DepMap_ID", "prism_auc_gossypol")), by = "DepMap_ID") %>%
  dplyr::mutate(prism_auc_gossypol = as.numeric(prism_auc_gossypol))

data_to_plot %<>% left_join(., prism_secondary_response %>% 
                              dplyr::filter(name == "BMS-387032") %>%
                              dplyr::select(depmap_id, auc) %>%
                              magrittr::set_colnames(c("DepMap_ID", "prism_auc_BMS387032")), by = "DepMap_ID") %>%
  dplyr::mutate(prism_auc_BMS387032 = as.numeric(prism_auc_BMS387032))

data_to_plot %<>% left_join(., prism_secondary_response %>% 
                              dplyr::filter(name == "morin") %>%
                              dplyr::select(depmap_id, auc) %>%
                              magrittr::set_colnames(c("DepMap_ID", "prism_auc_morin")), by = "DepMap_ID") %>%
  dplyr::mutate(prism_auc_morin = as.numeric(prism_auc_morin))

data_to_plot %<>% left_join(., sanger_dose_response %>% 
                              dplyr::filter(DRUG_NAME == "MIM1") %>%
                              dplyr::select(ARXSPAN_ID, AUC_PUBLISHED) %>%
                              magrittr::set_colnames(c("DepMap_ID", "sanger_auc_MIM1")), by = "DepMap_ID") %>%
  dplyr::mutate(sanger_auc_MIM1 = as.numeric(sanger_auc_MIM1))

data_to_plot %<>% left_join(., sanger_dose_response %>% 
                              dplyr::filter(DRUG_NAME == "TW-37") %>%
                              dplyr::select(ARXSPAN_ID, AUC_PUBLISHED) %>%
                              magrittr::set_colnames(c("DepMap_ID", "sanger_auc_TW37")), by = "DepMap_ID") %>%
  dplyr::mutate(sanger_auc_TW37 = as.numeric(sanger_auc_TW37))

data_to_plot %<>% left_join(., sanger_dose_response %>% 
                              dplyr::filter(DRUG_NAME == "OBATOCLAX") %>%
                              dplyr::select(ARXSPAN_ID, AUC_PUBLISHED) %>%
                              magrittr::set_colnames(c("DepMap_ID", "sanger_auc_obatoclax")), by = "DepMap_ID") %>%
  dplyr::mutate(sanger_auc_obatoclax = as.numeric(sanger_auc_obatoclax))

data_to_plot %<>% left_join(., sanger_dose_response %>% 
                              dplyr::filter(grepl("SABUTOCLAX", DRUG_NAME)) %>%
                              dplyr::select(ARXSPAN_ID, AUC_PUBLISHED) %>%
                              magrittr::set_colnames(c("DepMap_ID", "sanger_auc_sabutoclax")), by = "DepMap_ID") %>%
  dplyr::mutate(sanger_auc_sabutoclax = as.numeric(sanger_auc_sabutoclax))

data_to_plot %<>% left_join(., sanger_dose_response %>% 
                              dplyr::filter(grepl("UMI-77", DRUG_NAME)) %>%
                              dplyr::select(ARXSPAN_ID, AUC_PUBLISHED) %>%
                              magrittr::set_colnames(c("DepMap_ID", "sanger_auc_UMI77")), by = "DepMap_ID") %>%
  dplyr::mutate(sanger_auc_UMI77 = as.numeric(sanger_auc_UMI77))
```

### Comparing to other dependency datasets {.tabset .tabset-fade}

#### Sanger CRISPR

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(sanger_gene_effect)), 
                               aes(x = gene_effect, y = sanger_gene_effect, color = PvA == "Pediatric")) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('MCL1 gene effect (Sanger)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_sanger_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_sanger_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### DEMETER2 RNAi 

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(demeter_gene_effect)), 
                               aes(x = gene_effect, y = demeter_gene_effect, color = PvA == "Pediatric")) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('MCL1 gene effect (DEMETER2)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_demeter_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_demeter_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### DEMETER2 DRIVE RNAi 

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(demeter_drive_gene_effect)), 
                               aes(x = gene_effect, y = demeter_drive_gene_effect, color = PvA == "Pediatric")) + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('MCL1 gene effect (DEMETER2 DRIVE)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_demeter_drive_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_demeter_drive_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

### Comparing to drug datasets {.tabset .tabset-fade}

#### PRISM TW37

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(prism_auc_TW37)), 
                               aes(x = gene_effect, y = prism_auc_TW37, color = PvA == "Pediatric")) + 
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('TW-37 AUC (PRISM)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_prism_TW37_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_prism_TW37_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### PRISM morin

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(prism_auc_morin)), 
                               aes(x = gene_effect, y = prism_auc_morin, color = PvA == "Pediatric")) + 
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('Morin AUC (PRISM)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_prism_morin_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_prism_morin_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### PRISM gossypol

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(prism_auc_gossypol)), 
                               aes(x = gene_effect, y = prism_auc_gossypol, color = PvA == "Pediatric")) + 
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('Gossypol AUC (PRISM)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_prism_gossypol_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_prism_gossypol_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### PRISM BMS387032

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(prism_auc_BMS387032)), 
                               aes(x = gene_effect, y = prism_auc_BMS387032, color = PvA == "Pediatric")) + 
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('BMS-387032 AUC (PRISM)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_prism_BMS387032_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_prism_BMS387032_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### Sanger TW37

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(sanger_auc_TW37)), 
                               aes(x = gene_effect, y = sanger_auc_TW37, color = PvA == "Pediatric")) + 
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('TW-37 AUC (Sanger)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_sanger_TW37_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_sanger_TW37_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### Sanger MIM1

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(sanger_auc_MIM1)), 
                               aes(x = gene_effect, y = sanger_auc_MIM1, color = PvA == "Pediatric")) + 
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('MIM1 AUC (Sanger)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_sanger_MIM1_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_sanger_MIM1_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### Sanger UMI77

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(sanger_auc_UMI77)), 
                               aes(x = gene_effect, y = sanger_auc_UMI77, color = PvA == "Pediatric")) + 
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('UMI-77 AUC (Sanger)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_sanger_UMI77_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_sanger_UMI77_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### Sanger obatoclax

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(sanger_auc_obatoclax)), 
                               aes(x = gene_effect, y = sanger_auc_obatoclax, color = PvA == "Pediatric")) + 
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('Obatoclax AUC (Sanger)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_sanger_obatoclax_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_sanger_obatoclax_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### Sanger sabutoclax

```{r include=T}
MCL1_comparison_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(sanger_auc_sabutoclax)), 
                               aes(x = gene_effect, y = sanger_auc_sabutoclax, color = PvA == "Pediatric")) + 
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect (DepMap)') + ylab('Sabutoclax AUC (Sanger)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_sanger_sabutoclax_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_sanger_sabutoclax_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

### Comparing to heme malignancies {.tabset .tabset-fade}

#### Solid tumors to heme malignancies

```{r include=T}
load_gene_effect_heme <- function(){
  if(!file.exists("data/gene_effect.csv")) # If file does not exist, download from FigShare. Otherwise load local file.
  {
    gene_effect <- fread(gene_effect_url)
    fwrite(gene_effect, "data/gene_effect.csv")
  } else
  {
    gene_effect <- fread("data/gene_effect.csv")
  }
  # Reformat as a matrix.
  gene_effect %<>% dplyr::filter(!(V1 %in% mf$DepMap_ID))
  rownames_temp <- gene_effect$V1
  gene_effect <- as.matrix(gene_effect[,-1])
  rownames(gene_effect) <- rownames_temp
  return(gene_effect)
}

gene_effect_heme <- load_gene_effect_heme()

genes_to_highlight <- "MCL1 (4170)"

data_to_plot <- rbind(
  left_join(mf, 
            as.data.frame(cbind(rownames(gene_effect), gene_effect[,genes_to_highlight,drop=F])) %>% 
              magrittr::set_colnames(c("DepMap_ID", "gene_effect")), by = "DepMap_ID") %>%
    dplyr::select(DepMap_ID, PvA, gene_effect),
  as.data.frame(cbind(rownames(gene_effect_heme), gene_effect_heme[,genes_to_highlight,drop=F])) %>% 
    magrittr::set_colnames(c("DepMap_ID", "gene_effect")) %>%
    dplyr::mutate(PvA = "Heme") %>%
    dplyr::select(DepMap_ID, PvA, gene_effect)
  ) %>%
  dplyr::mutate(gene_effect = as.numeric(gene_effect)) %>%
  dplyr::filter(!is.na(PvA) & !is.na(gene_effect) & PvA != "Fibroblast")

### Plot cmparison of peds v adults v heme
data_to_plot$PvA <- factor(data_to_plot$PvA, levels = c("Pediatric", "Adult", "Heme"))
data_to_plot %<>% dplyr::group_by(PvA) %>%
  dplyr::mutate(n = n()) %>% 
  dplyr::mutate(label = paste0(PvA,' (n=',n,')')) %>%
  dplyr::ungroup()

MCL1_comparison_plot <- ggplot(data_to_plot, aes(x = PvA, y = gene_effect, fill=PvA, color=PvA)) +
  geom_violin(alpha=0.75) + 
  geom_boxplot(pch=21, fill = "white", color = "black", width = 0.1) + 
  scale_fill_manual(values=c('Pediatric'='red', 'Adult'='grey', 'Heme'='black')) +
  scale_color_manual(values=c('Pediatric'='red', 'Adult'='grey', 'Heme'='black')) +
  theme_bw() +
  xlab('') + ylab('MCL1 gene effect') +
  scale_x_discrete(labels=setNames(unique(data_to_plot$label), unique(data_to_plot$PvA))) +
  theme_Publication() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5), legend.position='none',
        panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  stat_compare_means(comparisons = list(c("Pediatric","Heme"), c("Pediatric","Adult"), c("Heme","Adult")), label.y = c(0.5, 0.1, 0.3)) +
  coord_cartesian(clip = "off")

MCL1_comparison_plot

saveRDS(MCL1_comparison_plot, file = paste0('figures/mcl1_compare/depmap_vs_heme_', version_to_use, '.rds'))
ggsave(MCL1_comparison_plot, filename = paste0('figures/mcl1_compare/depmap_vs_heme_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

## Session Info

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```

```{r include=TRUE}
as.data.frame(`Dataset Used`) %>% datatable(options=list(scrollX=T))
```
