---
title: "Predictive Features Figures"
author: "Neekesh Dharia & Guillaume Kugener"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, include=FALSE, message=FALSE, warning = FALSE)
source('setup.R')
source('load_data.R')
```


# Generated using: `r toupper(version_to_use)`


```{r load_needed_datasets}
# This script uses the following datasets:
mf <- mf
gene_dependency <- load_gene_dependency()
gene_effect <- load_gene_effect()
gene_expression <- load_gene_expression()
gene_cn <- load_gene_cn()
gene_lrt_table <- load_lrt_table()
achilles_common_essentials <- load_achilles_common_essentials()

if(!dir.exists('figures')) { dir.create('figures') }
if(!dir.exists('figures/biomarkers')) { dir.create('figures/biomarkers') }
```


The table below highlights all of the cell lines that are included in the analyses and the types that they are annotated as:

```{r}
cell_lines_included_table <- rbind(
  data.frame(name=row.names(gene_dependency), stringsAsFactors = F) %>% mutate(Source='Dependency'),
  data.frame(name=row.names(gene_expression), stringsAsFactors = F) %>% mutate(Source='Expression')
) %>% mutate(value=1) %>% mutate(DepMap_ID=name) %>%
  spread(., Source, value) %>%
  left_join(., mf %>% dplyr::select(DepMap_ID, CCLE_name, Type), by='DepMap_ID') %>%
  dplyr::select(DepMap_ID, CCLE_name, everything())
```

```{r include=TRUE} 
cell_lines_included_table %>% datatable(options=list(scrollX=T))
```

# Overview

* Evaluate predictive features for selective dependencies

```{r}
biomarkers <- fread("ensemble_prediction/Avana-solid-ensemble.csv")
biomarkers_peds <- fread("ensemble_prediction/Avana-pediatric-ensemble.csv")

biomarkers_top <- biomarkers %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(gene) %>% 
  dplyr::top_n(n = 1, pearson) %>%
  dplyr::ungroup() %>% 
  dplyr::mutate(percentile = rank(pearson) / 18333)
biomarkers_peds_top <- biomarkers_peds %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(gene) %>% 
  dplyr::top_n(n = 1, pearson) %>%
  dplyr::ungroup() %>% 
  dplyr::mutate(percentile = rank(pearson) / 18333)
```

```{r}
lrt_genes <- gene_lrt_table %>% filter((lrt) > 100, skewed_left) %$%
  gene

achilles_non_essentials <- colSums(gene_dependency >= 0.5, na.rm = T)
achilles_non_essentials <- names(achilles_non_essentials[achilles_non_essentials == 0])

lrt_genes_ceres <- lrt_genes
lrt_genes_ceres <- lrt_genes_ceres[!(lrt_genes_ceres %in% c(achilles_non_essentials, achilles_common_essentials$gene))]

dist_biomarkers <- ggplot(full_join(biomarkers_top, biomarkers_peds_top, 
                 by = "gene", suffix = c(".all", ".peds")) %>% 
         dplyr::filter(gene %in% lrt_genes_ceres), aes(pearson.peds)) +
  geom_density(color = "red") + 
  geom_density(data = full_join(biomarkers_top, biomarkers_peds_top, 
                                by = "gene", suffix = c(".all", ".peds")) %>% 
                 dplyr::filter(gene %in% lrt_genes_ceres), mapping = aes(pearson.all)) +
  theme_bw() +
  xlim(c(-.5, 1.0)) + 
  xlab('Pearson correlation of predictive model') + ylab('Density') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

saveRDS(dist_biomarkers, file = paste0('figures/biomarkers/pearson_dist_lrt_genes_', version_to_use, '.rds'))
ggsave(dist_biomarkers, filename = paste0('figures/biomarkers/pearson_dist_lrt_genes_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```


```{r}
gene_dependency_adult <- gene_dependency[rownames(gene_dependency) %in% mf$DepMap_ID[mf$PvA == "Adult"],]
gene_dependency_peds <- gene_dependency[rownames(gene_dependency) %in% mf$DepMap_ID[mf$PvA == "Pediatric"],]
gene_effect_adult <- gene_effect[rownames(gene_effect) %in% mf$DepMap_ID[mf$PvA == "Adult"],]
gene_effect_peds <- gene_effect[rownames(gene_effect) %in% mf$DepMap_ID[mf$PvA == "Pediatric"],]

# Ratio of cell lines dependent on selective dependecies
adult_lrt_rate <- colSums(gene_dependency_adult[, lrt_genes_ceres]>0.5, na.rm=T) / colSums(gene_dependency_adult[, lrt_genes_ceres]>=0, na.rm=T)
pediatric_lrt_rate <- colSums(gene_dependency_peds[, lrt_genes_ceres]>0.5, na.rm=T) / colSums(gene_dependency_peds[, lrt_genes_ceres]>=0, na.rm=T)
total_lrt_rate <- colSums(gene_dependency[, lrt_genes_ceres]>0.5, na.rm=T) / colSums(gene_dependency[, lrt_genes_ceres]>=0, na.rm=T)

diff_lrt_rate <- pediatric_lrt_rate - adult_lrt_rate

diff_lrt_rate_pvalue <- diff_lrt_rate

for (gene in lrt_genes_ceres)
{
  gene <- as.character(gene)
  diff_lrt_rate_pvalue[gene] <- fisher.test(matrix(c(sum(gene_dependency_adult[, gene]>0.5, na.rm=T),
                                                     sum(gene_dependency_adult[, gene]>=0, na.rm=T), 
                                                     sum(gene_dependency_peds[, gene]>0.5, na.rm=T),
                                                     sum(gene_dependency_peds[, gene]>=0, na.rm=T)), 
                                                   nrow = 2))$p.value
}
diff_lrt_rate_pvalue_adj <- p.adjust(diff_lrt_rate_pvalue, method = "BH")

all_lrt_rate_by_type <- gather(as.data.frame(gene_dependency[, lrt_genes_ceres]) %>% 
                                 dplyr::mutate(DepMap_ID = rownames(.)), Gene, dep, -DepMap_ID) %>%
  left_join(., mf %>% dplyr::select(DepMap_ID, Type, PvA), by='DepMap_ID') %>%
  group_by(Gene, Type) %>% 
  mutate(n = n()) %>%
  mutate(n_dep = sum(dep >= 0.5, na.rm=T)) %>%
  dplyr::select(Type, PvA, Gene, n, n_dep) %>%
  unique() %>%
  group_by(Gene, PvA) %>%
  mutate(lrt_rate = sum(n_dep) / sum (n), n_PvA = sum(n_dep)) %>%
  ungroup() %>%
  mutate(type_lrt_rate = lrt_rate * n_dep / n_PvA) %>%
  mutate(p.value = diff_lrt_rate_pvalue[Gene], p.value.adj = diff_lrt_rate_pvalue_adj[Gene]) 

all_lrt_rate_by_type %<>% rbind(., 
                                as.data.frame(cbind(names(diff_lrt_rate), diff_lrt_rate, diff_lrt_rate_pvalue, diff_lrt_rate_pvalue_adj)) %>% 
                                  dplyr::rename(Gene = V1, type_lrt_rate = diff_lrt_rate,
                                                p.value = diff_lrt_rate_pvalue, p.value.adj = diff_lrt_rate_pvalue_adj) %>% 
                                  mutate(Type = ifelse(as.numeric(type_lrt_rate) > 0, "Differential_pos", "Differential_neg"), 
                                         PvA = "Differential", 
                                         n = NA, n_dep = NA, lrt_rate = type_lrt_rate, n_PvA = NA) %>%
                                  dplyr::select((names(all_lrt_rate_by_type)))) %>%
  filter(PvA != "Fibroblast") %>%
  mutate(Group = ifelse(PvA == "Adult", "Fraction of\nadult lines\ndependent", 
                        ifelse(PvA == "Pediatric", "Fraction of\npediatric lines\ndependent", 
                               "Difference in\npediatric vs adult\nfraction"))) %>%
  mutate(Group = factor(Group, levels = c("Fraction of\nadult lines\ndependent", 
                                          "Fraction of\npediatric lines\ndependent", 
                                          "Difference in\npediatric vs adult\nfraction"))) %>%
  mutate(type_lrt_rate = as.numeric(type_lrt_rate), 
         lrt_rate = as.numeric(lrt_rate)) %>% 
  mutate(type_lrt_rate = ifelse(type_lrt_rate == "NaN", 0, type_lrt_rate)) %>%
  dplyr::left_join(., as.data.frame(cbind(names(total_lrt_rate), total_lrt_rate)) %>% magrittr::set_colnames(c("Gene", "total_lrt_rate")), by = "Gene") %>%
  dplyr::mutate(total_lrt_rate = as.numeric(total_lrt_rate))

```

```{r}
lrt_rates_and_biomarkers_top <- full_join(
  full_join(
    full_join(
      all_lrt_rate_by_type %>% 
        dplyr::select(Gene, PvA, lrt_rate, total_lrt_rate, p.value.adj) %>% 
        dplyr::filter(PvA != "Differential") %>% 
        unique(), 
      as.data.frame(cbind(names(diff_lrt_rate), diff_lrt_rate)) %>% magrittr::set_colnames(c("Gene", "diff_lrt_rate")), by = "Gene"), 
    biomarkers_top %>% 
      dplyr::rename(Gene = gene), by = "Gene"),
  biomarkers_peds_top %>% 
    dplyr::rename(Gene = gene), by = "Gene", suffix = c("", ".peds"))
```


### Predictive features for selective dependencies {.tabset .tabset-fade}

#### Dependency rate versus predictive Pearson

```{r include=T}
genes_to_highlight <- c("MCL1 (4170)", #"BCL2 (596)",
                        "CDK4 (1019)", #"CDK6 (1021)",
                        "MDM2 (4193)", #"MDM4 (4194)",
                        "ALK (238)", "BRAF (673)")

lrt_biomarkers_plot <- ggplot(lrt_rates_and_biomarkers_top %>% 
                                dplyr::filter(PvA == "Pediatric") %>%
                                dplyr::mutate(color_highlight = Gene %in% genes_to_highlight), 
                              aes(x = total_lrt_rate, y = pearson, 
                                  label = Gene, size = lrt_rate, 
                                  color = color_highlight)) + 
  geom_point(alpha = 0.5) +
  geom_text_repel(data = lrt_rates_and_biomarkers_top %>% 
                     dplyr::filter(PvA == "Pediatric") %>%
                     dplyr::filter(Gene %in% genes_to_highlight), 
                   size = 2, color = "black",
                   segment.size = 0.2, nudge_x = 0.05, nudge_y = 0.05) + 
  scale_size(name = "Rate of selective\ndependency in\npediatric lines", breaks = seq(0, 1, 0.2), range = c(0,3)) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('Rate of selective dependency') + ylab('Pearson correlation of predictive model') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

lrt_biomarkers_plot

saveRDS(lrt_biomarkers_plot, file = paste0('figures/biomarkers/lrt_rate_vs_biomarker_', version_to_use, '.rds'))
ggsave(lrt_biomarkers_plot, filename = paste0('figures/biomarkers/lrt_rate_vs_biomarker_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```


#### Pearson correlations - pediatric versus all

```{r include=T}
# genes_to_highlight <- lrt_rates_and_biomarkers_top %>% 
#   dplyr::filter(pearson.peds > 0.5, pearson < 0.5) %>%
#   dplyr::pull(Gene) %>%
#   unique() 

genes_to_highlight <- lrt_rates_and_biomarkers_top %>% 
  dplyr::filter(grepl("^TRIM8 ", Gene) | grepl("^MYOD1 ", Gene) | grepl("^HDAC2 ", Gene) | grepl("^IGF1R ", Gene) |
                  grepl("^CDK4 ", Gene) | grepl("^ALK ", Gene) | grepl("^MDM2 ", Gene) |
                  grepl("^MCL1 ", Gene) | grepl("^BRAF ", Gene)) %>%
  dplyr::pull(Gene) %>%
  unique() 

biomarkers_peds_vs_all_plot <- ggplot(lrt_rates_and_biomarkers_top %>% 
                                        dplyr::filter(PvA == "Pediatric"), 
                                      aes(x = pearson, y = pearson.peds, 
                                          label = Gene, size = -log10(as.numeric(p.value.adj)), 
                                          color = diff_lrt_rate > 0)) + 
  geom_point(alpha = 0.5) +
  # geom_hline(yintercept = 0.5, linetype = "dashed", color = "black") + 
  # geom_vline(xintercept = 0.5, linetype = "dashed", color = "black") + 
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  geom_label_repel(data = lrt_rates_and_biomarkers_top %>% 
                     dplyr::filter(PvA == "Pediatric") %>%
                     dplyr::filter(Gene %in% genes_to_highlight), 
                   size = 3, color = "black", fill = rgb(1, 1, 1, 0.5),
                   segment.size = 0.2, nudge_x = 0.05, nudge_y = 0.05) + 
  scale_size(name = "-log10\n(adjusted\np-value)") +
  scale_color_manual(values = c("gray", "red"), name = "Higher rate\nof dependency", labels = c("Adult", "Pediatric")) + 
  theme_bw() +
  xlab('Pearson correlation of predictive model\n(all cell lines)') + 
  ylab('Pearson correlation of predictive model\n(pediatric lines only)') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

biomarkers_peds_vs_all_plot

saveRDS(biomarkers_peds_vs_all_plot, file = paste0('figures/biomarkers/biomarker_peds_vs_all_', version_to_use, '.rds'))
ggsave(biomarkers_peds_vs_all_plot, filename = paste0('figures/biomarkers/biomarker_peds_vs_all_', version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```


### Example predictive models {.tabset .tabset-fade}

#### MCL1

```{r include=T, fig.height=6, fig.width=4}
genes_to_highlight <- lrt_rates_and_biomarkers_top %>% 
  dplyr::filter(grepl("^MCL1 ", Gene)) %>%
  dplyr::pull(Gene) %>%
  unique() 

gene_data <- lrt_rates_and_biomarkers_top %>% 
                                     dplyr::filter(PvA == "Pediatric") %>%
                                     dplyr::filter(Gene %in% genes_to_highlight) %>%
                                     dplyr::select(Gene, dplyr::starts_with("feature")) %>%
                                     tidyr::pivot_longer(cols = dplyr::contains("importance"), names_to = "cat", values_to = "importance") %>%
                                     tidyr::pivot_longer(cols = dplyr::starts_with("feature"), names_to = "feature_name", values_to = "feature") %>%
                                     dplyr::mutate(cat = gsub("_importance", "", cat)) %>%
                                     dplyr::filter(cat == feature_name) %>%
                                     dplyr::mutate(cat = ifelse(grepl(".peds", cat), "Peds", "All")) %>%
                                     dplyr::mutate(feature_plot = paste0(feature_name, ";", feature))

predictive_features_plot <- ggplot(gene_data %>% dplyr::filter(cat == "All"), 
                                   aes(x = feature_plot, y = importance)) + 
  geom_col(color = "gray", alpha = 0.5) +
  theme_bw() +
  scale_x_discrete(labels = function(x) { str_trunc(gsub(".*[;]", "", x), 30, side = "center", ellipsis = "...") }, name = "Feature") +
  ylab('Feature importance') +
  ggtitle(genes_to_highlight) +
  theme_Publication() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5), 
        panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
  
predictive_features_plot

saveRDS(predictive_features_plot, file = paste0('figures/biomarkers/pred_features_', genes_to_highlight, version_to_use, '.rds'))
ggsave(predictive_features_plot, filename = paste0('figures/biomarkers/pred_features_', genes_to_highlight, version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### IGF1R

```{r include=T, fig.height=6, fig.width=4}
genes_to_highlight <- lrt_rates_and_biomarkers_top %>% 
  dplyr::filter(grepl("^IGF1R ", Gene)) %>%
  dplyr::pull(Gene) %>%
  unique() 

gene_data <- lrt_rates_and_biomarkers_top %>% 
                                     dplyr::filter(PvA == "Pediatric") %>%
                                     dplyr::filter(Gene %in% genes_to_highlight) %>%
                                     dplyr::select(Gene, dplyr::starts_with("feature")) %>%
                                     tidyr::pivot_longer(cols = dplyr::contains("importance"), names_to = "cat", values_to = "importance") %>%
                                     tidyr::pivot_longer(cols = dplyr::starts_with("feature"), names_to = "feature_name", values_to = "feature") %>%
                                     dplyr::mutate(cat = gsub("_importance", "", cat)) %>%
                                     dplyr::filter(cat == feature_name) %>%
                                     dplyr::mutate(cat = ifelse(grepl(".peds", cat), "Peds", "All")) %>%
                                     dplyr::mutate(feature_plot = paste0(feature_name, ";", feature))

predictive_features_plot <- ggplot(gene_data %>% dplyr::filter(cat == "All"), 
                                   aes(x = feature_plot, y = importance)) + 
  geom_col(color = "gray", alpha = 0.5) +
  theme_bw() +
  scale_x_discrete(labels = function(x) { str_trunc(gsub(".*[;]", "", x), 30, side = "center", ellipsis = "...") }, name = "Feature") +
  ylab('Feature importance') +
  ggtitle(genes_to_highlight) +
  theme_Publication() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5), 
        panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
  
predictive_features_plot


predictive_features_plot <- ggplot(gene_data, 
                                   aes(x = feature_plot, y = importance)) + 
  geom_col(color = "gray", alpha = 0.5) +
  theme_bw() +
  scale_x_discrete(labels = function(x) { str_trunc(gsub(".*[;]", "", x), 30, side = "center", ellipsis = "...") }, name = "Feature") +
  ylab('Feature importance') +
  ggtitle(genes_to_highlight) +
  theme_Publication() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5), 
        panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  facet_grid(~cat, scales = "free")
  

predictive_features_plot

saveRDS(predictive_features_plot, file = paste0('figures/biomarkers/pred_features_', genes_to_highlight, version_to_use, '.rds'))
ggsave(predictive_features_plot, filename = paste0('figures/biomarkers/pred_features_', genes_to_highlight, version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

#### HDAC2

```{r include=T, fig.height=6, fig.width=4}
genes_to_highlight <- lrt_rates_and_biomarkers_top %>% 
  dplyr::filter(grepl("^HDAC2 ", Gene)) %>%
  dplyr::pull(Gene) %>%
  unique() 

gene_data <- lrt_rates_and_biomarkers_top %>% 
                                     dplyr::filter(PvA == "Pediatric") %>%
                                     dplyr::filter(Gene %in% genes_to_highlight) %>%
                                     dplyr::select(Gene, dplyr::starts_with("feature")) %>%
                                     tidyr::pivot_longer(cols = dplyr::contains("importance"), names_to = "cat", values_to = "importance") %>%
                                     tidyr::pivot_longer(cols = dplyr::starts_with("feature"), names_to = "feature_name", values_to = "feature") %>%
                                     dplyr::mutate(cat = gsub("_importance", "", cat)) %>%
                                     dplyr::filter(cat == feature_name) %>%
                                     dplyr::mutate(cat = ifelse(grepl(".peds", cat), "Peds", "All")) %>%
                                     dplyr::mutate(feature_plot = paste0(feature_name, ";", feature))

predictive_features_plot <- ggplot(gene_data %>% dplyr::filter(cat == "All"), 
                                   aes(x = feature_plot, y = importance)) + 
  geom_col(color = "gray", alpha = 0.5) +
  theme_bw() +
  scale_x_discrete(labels = function(x) { str_trunc(gsub(".*[;]", "", x), 30, side = "center", ellipsis = "...") }, name = "Feature") +
  ylab('Feature importance') +
  ggtitle(genes_to_highlight) +
  theme_Publication() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5), 
        panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
  
predictive_features_plot


predictive_features_plot <- ggplot(gene_data, 
                                   aes(x = feature_plot, y = importance)) + 
  geom_col(color = "gray", alpha = 0.5) +
  theme_bw() +
  scale_x_discrete(labels = function(x) { str_trunc(gsub(".*[;]", "", x), 30, side = "center", ellipsis = "...") }, name = "Feature") +
  ylab('Feature importance') +
  ggtitle(genes_to_highlight) +
  theme_Publication() +
  theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5), 
        panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +
  facet_grid(~cat, scales = "free")
  

predictive_features_plot

saveRDS(predictive_features_plot, file = paste0('figures/biomarkers/pred_features_', genes_to_highlight, version_to_use, '.rds'))
ggsave(predictive_features_plot, filename = paste0('figures/biomarkers/pred_features_', genes_to_highlight, version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```


### Example predictive model scatter plots {.tabset .tabset-fade}

#### MCL1 - BCL2L1

```{r include=T}
genes_to_highlight <- "MCL1_BCL2L1"

data_to_plot <- left_join(
  left_join(mf, 
            as.data.frame(cbind(rownames(gene_effect), gene_effect[,"MCL1 (4170)",drop=F])) %>% 
              set_colnames(c("DepMap_ID", "gene_effect")), by = "DepMap_ID"),
  as.data.frame(cbind(rownames(gene_expression), gene_expression[,"BCL2L1 (598)",drop=F])) %>% 
    set_colnames(c("DepMap_ID", "gene_expression")), by = "DepMap_ID") %>%
  dplyr::mutate(gene_effect = as.numeric(gene_effect), 
                gene_expression = as.numeric(gene_expression))
  

biomarker_scatter_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(gene_expression)), 
                              aes(x = gene_effect, y = gene_expression, color = PvA == "Pediatric")) + 
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) + 
  theme_bw() +
  xlab('MCL1 gene effect') + ylab('BCL2L1 gene expression (log2(TPM))') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

biomarker_scatter_plot

saveRDS(biomarker_scatter_plot, file = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.rds'))
ggsave(biomarker_scatter_plot, filename = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.pdf'), 
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```


<!-- #### MDM2 - EDA2R -->

<!-- ```{r include=T} -->
<!-- genes_to_highlight <- "MDM2_EDA2R" -->

<!-- data_to_plot <- left_join( -->
<!--   left_join(mf,  -->
<!--             as.data.frame(cbind(rownames(gene_effect), gene_effect[,"MDM2 (4193)",drop=F])) %>%  -->
<!--               set_colnames(c("DepMap_ID", "gene_effect")), by = "DepMap_ID"), -->
<!--   as.data.frame(cbind(rownames(gene_expression), gene_expression[,"EDA2R (60401)",drop=F])) %>%  -->
<!--     set_colnames(c("DepMap_ID", "gene_expression")), by = "DepMap_ID") %>% -->
<!--   dplyr::mutate(gene_effect = as.numeric(gene_effect),  -->
<!--                 gene_expression = as.numeric(gene_expression)) -->


<!-- biomarker_scatter_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(gene_expression)),  -->
<!--                               aes(x = gene_effect, y = gene_expression, color = PvA == "Pediatric")) +  -->
<!--   stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) + -->
<!--   geom_point(alpha = 0.5) + -->
<!--   scale_color_manual(values = c("gray", "red"), guide = F) +  -->
<!--   theme_bw() + -->
<!--   xlab('MDM2 gene effect') + ylab('EDA2R gene expression (log2(TPM))') + -->
<!--   theme_Publication() + -->
<!--   theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1, -->
<!--         panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))  -->

<!-- biomarker_scatter_plot -->

<!-- saveRDS(biomarker_scatter_plot, file = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.rds')) -->
<!-- ggsave(biomarker_scatter_plot, filename = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.pdf'),  -->
<!--        height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE) -->
<!-- ``` -->

<!-- #### HDAC2 - FUCA1 -->

<!-- ```{r include=T} -->
<!-- genes_to_highlight <- "HDAC2_FUCA1" -->

<!-- data_to_plot <- left_join( -->
<!--   left_join(mf,  -->
<!--             as.data.frame(cbind(rownames(gene_effect), gene_effect[,"HDAC2 (3066)",drop=F])) %>%  -->
<!--               set_colnames(c("DepMap_ID", "gene_effect")), by = "DepMap_ID"), -->
<!--   as.data.frame(cbind(rownames(gene_expression), gene_expression[,"FUCA1 (2517)",drop=F])) %>%  -->
<!--     set_colnames(c("DepMap_ID", "gene_expression")), by = "DepMap_ID") %>% -->
<!--   dplyr::mutate(gene_effect = as.numeric(gene_effect),  -->
<!--                 gene_expression = as.numeric(gene_expression)) -->


<!-- biomarker_scatter_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(gene_expression)),  -->
<!--                               aes(x = gene_effect, y = gene_expression, color = PvA == "Pediatric")) +  -->
<!--   stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) + -->
<!--   geom_point(alpha = 0.5) + -->
<!--   scale_color_manual(values = c("gray", "red"), guide = F) +  -->
<!--   theme_bw() + -->
<!--   xlab('HDAC2 gene effect') + ylab('FUCA1 gene expression (log2(TPM))') + -->
<!--   theme_Publication() + -->
<!--   theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1, -->
<!--         panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))  -->

<!-- biomarker_scatter_plot -->

<!-- saveRDS(biomarker_scatter_plot, file = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.rds')) -->
<!-- ggsave(biomarker_scatter_plot, filename = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.pdf'),  -->
<!--        height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE) -->
<!-- ``` -->

#### IGF1R - IGF1R

```{r include=T}
genes_to_highlight <- "IGF1R_IGF1R"

data_to_plot <- left_join(
  left_join(mf,
            as.data.frame(cbind(rownames(gene_effect), gene_effect[,"IGF1R (3480)",drop=F])) %>%
              set_colnames(c("DepMap_ID", "gene_effect")), by = "DepMap_ID"),
  as.data.frame(cbind(rownames(gene_expression), gene_expression[,"IGF1R (3480)",drop=F])) %>%
    set_colnames(c("DepMap_ID", "gene_expression")), by = "DepMap_ID") %>%
  dplyr::mutate(gene_effect = as.numeric(gene_effect),
                gene_expression = as.numeric(gene_expression))


biomarker_scatter_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(gene_expression)),
                              aes(x = gene_effect, y = gene_expression, color = PvA == "Pediatric")) +
  stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("gray", "red"), guide = F) +
  theme_bw() +
  xlab('IGF1R gene effect') + ylab('IGF1R gene expression (log2(TPM))') +
  theme_Publication() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1,
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

biomarker_scatter_plot

saveRDS(biomarker_scatter_plot, file = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.rds'))
ggsave(biomarker_scatter_plot, filename = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.pdf'),
       height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE)
```

<!-- #### IGF1R - IGF2 -->

<!-- ```{r include=T} -->
<!-- genes_to_highlight <- "IGF1R_IGF2" -->

<!-- data_to_plot <- left_join( -->
<!--   left_join(mf,  -->
<!--             as.data.frame(cbind(rownames(gene_effect), gene_effect[,"IGF1R (3480)",drop=F])) %>%  -->
<!--               set_colnames(c("DepMap_ID", "gene_effect")), by = "DepMap_ID"), -->
<!--   as.data.frame(cbind(rownames(gene_expression), gene_expression[,"IGF2 (3481)",drop=F])) %>%  -->
<!--     set_colnames(c("DepMap_ID", "gene_expression")), by = "DepMap_ID") %>% -->
<!--   dplyr::mutate(gene_effect = as.numeric(gene_effect),  -->
<!--                 gene_expression = as.numeric(gene_expression)) -->


<!-- biomarker_scatter_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(gene_expression)),  -->
<!--                               aes(x = gene_effect, y = gene_expression, color = PvA == "Pediatric")) +  -->
<!--   stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) + -->
<!--   geom_point(alpha = 0.5) + -->
<!--   scale_color_manual(values = c("gray", "red"), guide = F) +  -->
<!--   theme_bw() + -->
<!--   xlab('IGF1R gene effect') + ylab('IGF2 gene expression (log2(TPM))') + -->
<!--   theme_Publication() + -->
<!--   theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1, -->
<!--         panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))  -->

<!-- biomarker_scatter_plot -->

<!-- saveRDS(biomarker_scatter_plot, file = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.rds')) -->
<!-- ggsave(biomarker_scatter_plot, filename = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.pdf'),  -->
<!--        height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE) -->
<!-- ``` -->

<!-- ### Example predictive model scatter plots with gene-sets {.tabset .tabset-fade} -->

<!-- #### IGF1R - "IGF" gene sets -->

<!-- ```{r include=T} -->
<!-- msigdb <- load_msigdb() -->

<!-- gene_sigs <- as.data.frame(unique(msigdb$C2$term), stringsAsFactors = F) %>%  -->
<!--   magrittr::set_colnames("V1") %>% -->
<!--   dplyr::filter(grepl("IGF", V1) & (grepl("BIOCARTA", V1) | grepl("REACTOME", V1))) -->

<!-- for (sig in gene_sigs$V1) -->
<!-- { -->
<!--   genes_to_highlight <- paste0("IGF1R_", sig) -->
<!--   sig_gene_members <- msigdb$C2$gene[msigdb$C2$term == sig] -->

<!--   data_to_plot <- left_join( -->
<!--     left_join(mf,  -->
<!--               as.data.frame(cbind(rownames(gene_effect), gene_effect[,"IGF1R (3480)",drop=F])) %>%  -->
<!--                 set_colnames(c("DepMap_ID", "gene_effect")), by = "DepMap_ID"), -->
<!--     as.data.frame(cbind(rownames(gene_expression),  -->
<!--                         rowMeans(gene_expression[,gsub("[)]", "", gsub(".* [(]", "",colnames(gene_expression))) %in% sig_gene_members,drop=F]))) %>%  -->
<!--       set_colnames(c("DepMap_ID", "sig_mean_expression")), by = "DepMap_ID") %>% -->
<!--     dplyr::mutate(gene_effect = as.numeric(gene_effect),  -->
<!--                   sig_mean_expression = as.numeric(sig_mean_expression)) -->

<!--   biomarker_scatter_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(sig_mean_expression)),  -->
<!--                                    aes(x = gene_effect, y = sig_mean_expression, color = PvA == "Pediatric")) +  -->
<!--     stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) + -->
<!--     geom_point(alpha = 0.5) + -->
<!--     scale_color_manual(values = c("gray", "red"), guide = F) +  -->
<!--     theme_bw() + -->
<!--     xlab('IGF1R gene effect') + ylab(paste0(str_trunc(sig, 30, side = "center", ellipsis = "..."), '\nmean gene expression (log2(TPM))')) + -->
<!--     theme_Publication() + -->
<!--     theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1, -->
<!--           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))  -->

<!--   print(biomarker_scatter_plot) -->

<!--   saveRDS(biomarker_scatter_plot, file = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.rds')) -->
<!--   ggsave(biomarker_scatter_plot, filename = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.pdf'),  -->
<!--          height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE) -->
<!-- } -->
<!-- ``` -->

<!-- #### MDM2 - "TP53" gene sets -->

<!-- ```{r include=T} -->
<!-- msigdb <- load_msigdb() -->

<!-- gene_sigs <- as.data.frame(unique(msigdb$C2$term), stringsAsFactors = F) %>%  -->
<!--   magrittr::set_colnames("V1") %>% -->
<!--   dplyr::filter(grepl("TP53_TARGETS", V1)) -->

<!-- for (sig in gene_sigs$V1) -->
<!-- { -->
<!--   genes_to_highlight <- paste0("MDM2_", sig) -->
<!--   sig_gene_members <- msigdb$C2$gene[msigdb$C2$term == sig] -->

<!--   data_to_plot <- left_join( -->
<!--     left_join(mf,  -->
<!--               as.data.frame(cbind(rownames(gene_effect), gene_effect[,"MDM2 (4193)",drop=F])) %>%  -->
<!--                 set_colnames(c("DepMap_ID", "gene_effect")), by = "DepMap_ID"), -->
<!--     as.data.frame(cbind(rownames(gene_expression),  -->
<!--                         rowMeans(gene_expression[,gsub("[)]", "", gsub(".* [(]", "",colnames(gene_expression))) %in% sig_gene_members,drop=F]))) %>%  -->
<!--       set_colnames(c("DepMap_ID", "sig_mean_expression")), by = "DepMap_ID") %>% -->
<!--     dplyr::mutate(gene_effect = as.numeric(gene_effect),  -->
<!--                   sig_mean_expression = as.numeric(sig_mean_expression)) -->

<!--   biomarker_scatter_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(sig_mean_expression)),  -->
<!--                                    aes(x = gene_effect, y = sig_mean_expression, color = PvA == "Pediatric")) +  -->
<!--     stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) + -->
<!--     geom_point(alpha = 0.5) + -->
<!--     scale_color_manual(values = c("gray", "red"), guide = F) +  -->
<!--     theme_bw() + -->
<!--     xlab('MDM2 gene effect') + ylab(paste0(str_trunc(sig, 30, side = "center", ellipsis = "..."), '\nmean gene expression (log2(TPM))')) + -->
<!--     theme_Publication() + -->
<!--     theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1, -->
<!--           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))  -->

<!--   print(biomarker_scatter_plot) -->

<!--   saveRDS(biomarker_scatter_plot, file = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.rds')) -->
<!--   ggsave(biomarker_scatter_plot, filename = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.pdf'),  -->
<!--          height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE) -->
<!-- } -->
<!-- ``` -->

<!-- #### HDAC2 - "HDAC2" gene sets -->

<!-- ```{r include=T} -->
<!-- msigdb <- load_msigdb() -->

<!-- gene_sigs <- as.data.frame(unique(msigdb$C2$term), stringsAsFactors = F) %>%  -->
<!--   magrittr::set_colnames("V1") %>% -->
<!--   dplyr::filter(grepl("HDAC2", V1)) -->

<!-- for (sig in gene_sigs$V1) -->
<!-- { -->
<!--   genes_to_highlight <- paste0("HDAC2_", sig) -->
<!--   sig_gene_members <- msigdb$C2$gene[msigdb$C2$term == sig] -->

<!--   data_to_plot <- left_join( -->
<!--     left_join(mf,  -->
<!--               as.data.frame(cbind(rownames(gene_effect), gene_effect[,"HDAC2 (3066)",drop=F])) %>%  -->
<!--                 set_colnames(c("DepMap_ID", "gene_effect")), by = "DepMap_ID"), -->
<!--     as.data.frame(cbind(rownames(gene_expression),  -->
<!--                         rowMeans(gene_expression[,gsub("[)]", "", gsub(".* [(]", "",colnames(gene_expression))) %in% sig_gene_members,drop=F]))) %>%  -->
<!--       set_colnames(c("DepMap_ID", "sig_mean_expression")), by = "DepMap_ID") %>% -->
<!--     dplyr::mutate(gene_effect = as.numeric(gene_effect),  -->
<!--                   sig_mean_expression = as.numeric(sig_mean_expression)) -->

<!--   biomarker_scatter_plot <- ggplot(data_to_plot %>% filter(!is.na(gene_effect), !is.na(sig_mean_expression)),  -->
<!--                                    aes(x = gene_effect, y = sig_mean_expression, color = PvA == "Pediatric")) +  -->
<!--     stat_smooth(method = "lm", formula = "y ~ x", alpha = 0.2) + -->
<!--     geom_point(alpha = 0.5) + -->
<!--     scale_color_manual(values = c("gray", "red"), guide = F) +  -->
<!--     theme_bw() + -->
<!--     xlab('HDAC2 gene effect') + ylab(paste0(str_trunc(sig, 30, side = "center", ellipsis = "..."), '\nmean gene expression (log2(TPM))')) + -->
<!--     theme_Publication() + -->
<!--     theme(panel.border = element_blank(), panel.grid.major = element_blank(), aspect.ratio = 1, -->
<!--           panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))  -->

<!--   print(biomarker_scatter_plot) -->

<!--   saveRDS(biomarker_scatter_plot, file = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.rds')) -->
<!--   ggsave(biomarker_scatter_plot, filename = paste0('figures/biomarkers/pred_features_scatter_', genes_to_highlight, version_to_use, '.pdf'),  -->
<!--          height = 14, width=15, units='cm', device='pdf', useDingbats=FALSE) -->
<!-- } -->
<!-- ``` -->

## Session Info

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```

```{r include=TRUE}
as.data.frame(`Dataset Used`) %>% datatable(options=list(scrollX=T))
```
